Geographically aware estimates of remotely sensed water properties for Chesapeake Bay

### 2022-09-26 manuscript/manuscript.tex
text edits

To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity at daily time-steps (over decadal time scales) as required by next-generation ESMs. The primary input to our model is remotely sensed reflectance data. In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in coastal Louisiana \citep{wangDevelopmentMODISData2018}. We supplemented MODIS data with gridded estimates of tributary inflows which we used to enforce physical realism (i.e. geographic “awareness”) especially in areas of the Bay with little to no observations. Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g. \citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.

@@

Because Chesapeake Bay has a mean optical depth of approximately 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column. We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours). Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}. This annual cycle was added back to model estimates for evaluation purposes and in the production of prediction surfaces. As a final observational data processing step, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.

>>>>

To address this validation data gap, we developed a data-driven model to produce high-resolution (< 1 $km^2$) estimates of temperature, salinity, and turbidity at daily time-steps (over decadal time scales) as required by next-generation ESMs. The primary input to our model is remotely sensed reflectance data. In particular, we use Moderate Resolution Imaging Spectroradiometer (MODIS) Aqua multiband data, which has been shown to be predictive of salinity in the Bay as well as in coastal Louisiana \citep{vogelAssessingSatelliteSea2016, wangDevelopmentMODISData2018}. We supplemented MODIS data with gridded estimates of tributary inflows which we used to enforce physical realism (i.e. geographic “awareness”) especially in areas of the Bay with little to no observations. Our model represents a further advantage over prior efforts at data-driven modeling of salinity in the Bay (e.g. \citet{vogelAssessingSatelliteSea2016, murphyComparisonSpatialInterpolation2010}) because it is calibrated against not only observations taken in the mainstem of the Bay but also observations taken in the mid to lower reaches of Bay tributaries.

@@

Because Chesapeake Bay has a mean optical depth of approximately 0.89m \citep{urquhartRemotelySensedEstimates2012}, we filtered all collected observations to those occurring in the top 1m of the water column. We further limited observations to those that had a timestamp during the MODIS overpass window time-of-day (approximately 1700 - 1900 hours GMT). Prior to model fitting, we removed the annual cycle from temperature observations using least squares optimization routines \citep{virtanenSciPyFundamentalAlgorithms2020a}. This annual cycle was added back to model estimates for evaluation purposes and in the production of prediction surfaces. As a final observational data processing step, we implemented spatial and temporal averaging of observational data to a daily mean within each MODIS pixel.
